---
title: "Visualization_Homework"
output: github_document
---

```{r setup, echo=TRUE, results='hide'}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

#### Load in data 
There are 1384617 observations of 15 variables in total. It contains order details from instacart such as order id, user id, order hour, product name, product department e.t.c . There are 39123 distinct products, 131209 order id and 131209 users id in the table. 
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
instacart %>% 
  count(product_name)

instacart %>% 
  count(user_id)

instacart %>% 
  count(order_id)
```

#### Number of aisles and most popular aisles.

There are 134 aisles. The top three popular aisles are fresh vegetables, fresh fruits, packaged vegetables fruits.
```{r}
instacart %>% 
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))
```

#### Plot showing the number of items ordered in each aisle
```{r}
instacart %>% 
  group_by(aisle_id, aisle) %>%
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>%
  ggplot(aes(x = aisle_id, y = n_obs, color = aisle)) +
  geom_point() +
  labs(
    title = "Aisles v.s Order number",
    x = "Aisle_id",
    y = "Number of items ordered"
  ) 
```


#### Table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. 

```{r}
instacart %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_obs = n()) %>% 
  filter(aisle == "baking ingredients" | aisle =="dog food care" |aisle == "packaged vegetables fruits") %>% 
  mutate(rank = min_rank(desc(n_obs))) %>% 
  filter(rank < 4)
```

#### Table showing mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day
```{r}
instacart %>% 
    mutate(order_hour_of_day= as.integer(order_hour_of_day)) %>% 
    select(product_name, order_dow, order_hour_of_day) %>% 
    filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
    group_by(product_name, order_dow) %>% 
    summarize(mean_order = mean(order_hour_of_day)) %>% 
    pivot_wider(
    names_from = order_dow,
    values_from = mean_order) 
```

## Problem 2

I cleaned the column names, created a column indicating "weekday" and "weekend", and perform pivot longer to "activity." columns. Now I have 6 colunms - "week", "day", "day_id", "day_of_week", "hour", "counts". There's 50400 observations of 6 variables. 
```{r}
accel_df =
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(day_of_week = ifelse(day == "Saturday", "weekend", ifelse(day == "Sunday", "weekend", "weekday"))) %>% 
  pivot_longer(
    starts_with("activity"),
    names_to = "minute",
    values_to = "counts"
  )
```

#### Total activity variable for each day

For week 1,2. There's increasing trend of summation of activity counts from Monday to Sunday.

```{r}
accel_df %>% 
  group_by(week, day) %>% 
  summarize(sum = sum(counts)) %>% 
  arrange(week, sum)
```
 
#### Graph presenting activity counts for each day 
Trends of graph: Most of Wednesday's and Tuesday's activity counts are low compared to counts on other days. There's also trend of one week of high counts followed by one week of low counts.
```{r}
accel_df %>% 
  ggplot(aes(x = minute, y = counts, color = day)) +
  geom_point(alpha = 0.5) +
    labs(
    title = "Minute v.s Activity Counts",
    x = "Minute",
    y = "Counts of Activity"
  ) 
```


## Problem 3

Data load in and description
```{r}
library(p8105.datasets)
data("ny_noaa")
```

